---
title: "Simulations of estimating causal effect curve using undersmoothed-HAL"
author: "Seraphina Shi"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r load_lib, include = FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(data.table)
library(dplyr)
library(tidyr)
library(foreach)
library(stringr)
library(glmnet)
library(sandwich)
library(ggplot2)
library(gridExtra)
library(grid)
```


# Data simulation 
Data structure:  $O = (W, A, Z, Y)$ 

 * U - exogenous variables  
 * W - baseline covariate that is a measure of body condition  
 * A - treatment level based on W, continuous between 0 and 5  
 * Y - outcome, indicator of an event ?  
   
 Underlying data generating process, $P_{U,X}$
 
* Exogenous variables:  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 1^2)$  
  + $U_A \sim Normal(\mu=0, \sigma^2 = 2^2)$  
  + $U_Y \sim Uniform(min = 0, max = 1)$  
  
* Structural equations F and endogenous variables:  
  + $W =  U_W$  
  + $A = bound(2 - 0.5W + U_A, min=0, max=5)$  
  + $Y = \mathbf{I}[U_Y < expit(W + 5*A  - 8)]$


```{r}
generate_data_1 <- function(n, a=NA){
  # exogenous variables
  U_W <- rnorm(n, 0, 1)
  U_A <- rnorm(n, 0, 2)
  U_Y <- runif(n, 0, 1)
  
  # endogenous variables
  W <- U_W
  
  if(is.na(a)){
    A <-  2 - 0.5*W + U_A
    A[A<=0] = 0
    A[A>=5] = 5
  } else {
    A <- rep(a, n)
  }
  
  Y <- as.numeric(U_Y < plogis(W + 5*A  - 8))
  
  # data frame
  O <- data.frame(W, A, Y)
  return(O)
}
obs <- generate_data_1(n=10000)
print(summary(obs))
```

# Calculate IC
```{r}
cal_IC_for_beta <- function(X, Y, Y_hat, beta_n){
  n <- dim(X)[1] 
  p <- length(beta_n)
  if (!is.matrix(X)) X <- as.matrix(X)
  
  # 1. calculate score: X(Y - phi(X))
  res <- Y-Y_hat
  score <- sweep(X, 1, res, `*`)
  
  # 2. calculate the derivative of phi:
  d_phi_scaler <- as.vector(exp(- beta_n %*% t(X)) / ((1 + exp(- beta_n %*% t(X)))^2))
  d_phi <- sweep(X, 1, d_phi_scaler, `*`)
  
  # 3. E_{P_n}(X d_phi)^(-1)
  tmat <- solve(t(X) %*% d_phi / n)
  
  # 4. calculate influence curves
  IC <- t(tmat %*% t(score))
  
  return(IC)
}
cal_IC_for_phi <- function(X_new, beta_n, IC_beta){
  
  if (!is.matrix(X_new)) X_new <- as.matrix(X_new)
  
  d_phi_scaler_new <- as.vector(exp(- beta_n %*% t(X_new)) / ((1 + exp(- beta_n %*% t(X_new)))^2))
  d_phi_new <- sweep(X_new, 1, d_phi_scaler_new, `*`)
  
  IC = diag(d_phi_new %*% t(IC_beta))
  
  return(IC)
}
```

```{r}
set.seed(123)
n = 100
obs <- generate_data_1(n)
glm_fit <- glm(formula = Y ~ W + A,
               family = binomial,
               data = obs)
y_name = "Y" 
x_names = c("W", "A")
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
  
  
ICs <- cal_IC_for_beta(X = cbind(rep(1, n), X),
                       Y = Y, 
                       Y_hat = predict(glm_fit, type = "response"), 
                       beta_n = glm_fit$coef)
se_IC <- sqrt(apply(ICs, 2, var)/n)
# compare estimated se from calculated ICs and from glm_fit. 
cbind(se_IC, sqrt(diag(sandwich(glm_fit))))
```

# Calculate IC for phi
```{r}
y_name = "Y" 
x_names = c("W", "A")
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
beta_n = glm_fit$coef
IC_beta <- cal_IC_for_beta(X = cbind(rep(1, n), X), Y = Y,
                           Y_hat = predict(glm_fit, type = "response"), 
                           beta_n = beta_n)
IC_phi <- cal_IC_for_phi(X_new  = cbind(rep(1, n), X), 
                         beta_n, IC_beta)
  
se_IC <- sqrt(var(IC_phi)/n)
se_Y_hat <- sqrt(var(predict(glm_fit))/n)
cbind(se_IC, se_Y_hat)
```

# Calculate IC for phi with HAL working model
```{r}
library(hal9001)
y_name = "Y" 
x_names = c("W", "A")
Y <- as.numeric(as.matrix(obs %>% select(all_of(y_name))))
X <- obs %>% 
    select(all_of(x_names)) %>% 
    mutate_if(sapply(., is.factor), as.numeric)
fit_init <- fit_hal(X = X,
                      Y = Y,
                      smoothness_orders = 0,
                      return_x_basis = TRUE,
                      family = "binomial",
                      num_knots = hal9001:::num_knots_generator(
                        max_degree = ifelse(ncol(X) >= 20, 2, 3),
                        smoothness_orders = 0,
                        base_num_knots_0 = max(100, ceiling(sqrt(n)))
                      )
)
Y_hat_init <- predict(fit_init, new_data = X)
init_coef <- fit_init$coefs[-1]
nonzero_col <- which(init_coef != 0)
init_coef_nonzero <- init_coef[nonzero_col]
basis_mat <- as.matrix(fit_init$x_basis)
basis_mat <- as.matrix(basis_mat[, nonzero_col])
IC_beta <- cal_IC_for_beta(X = basis_mat, Y = Y, Y_hat = Y_hat_init, 
                           beta_n = init_coef_nonzero)
IC_phi <- cal_IC_for_phi(X_new = basis_mat, 
                         beta_n = init_coef_nonzero, IC_beta)
se_IC <- sqrt(var(IC_phi)/n)
se_IC
```

```{r}
X_a1 <- X
X_a1$A = 1
Y_hat_init_a1 <- predict(fit_init, new_data = X_a1)
x_basis_a1 <- make_design_matrix(as.matrix(X_a1), fit_init$basis_list, p_reserve = 0.75)
x_basis_a1 <- as.matrix(x_basis_a1[, nonzero_col])
IC_phi <- cal_IC_for_phi(X_new = x_basis_a1, 
                         beta_n = init_coef_nonzero, IC_beta)
# matrix is singular and cannot be inverted
se_IC <- sqrt(var(IC_phi)/n)
se_IC
```



